---
title: "rwanda_ecd"
output: html_document
date: "2025-01-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Importing data and libraries
```{r}
# Load libraries
library(tidyverse)
library(here)
library(viridis)
library(sf)
library(tmap)
library(ggpp)
library(raster)
library(rasterVis)
library(readxl)
library(xml2)
library("RColorBrewer")
library(foreign)
library(haven)
library(Hmisc)
library(fuzzyjoin)

```


```{r}
#Load the rwanda survey files

rwa_household <- read_dta("data/cs_S0_S5_Household.dta")
rwa_individual <- read_dta("data/cs_S1_S2_S3_S4_S6A_S6E_Person.dta")
rwa_exp_annual <- read_dta("data/cs_S8A1_expenditure.dta")
rwa_exp_monthly <- read_dta("data/cs_S8A2_expenditure.dta")
rwa_exp_weekly <- read_dta("data/cs_S8A3_expenditure.dta")
rwa_exp_weekly2 <- read_dta("data/cs_S8B_expenditure.dta")
rwa_exp_other <- read_dta("data/cs_S9E_Other_expenditure.dta")
rwa_inc <- read_dta("data/cs_S9D_other_income.dta")
```


```{r}
#rds file to convert rwandan shillings to usd-ppp in 2025

ppp_2017 <- read_rds("data/01_010_conversion_factors.rds")
```

## Data cleaning
```{r}
#Convert select columns to factor columns to remove encoding

#Columns to convert to factors

factor_columns <- c("province", "district", "ur", "region", "poverty", "quintile" ,"s8a1q0", "s8a2q0", "s8a3q0", "s8a1q1", "s8a2q1", "s8a3q1", "s8a1q2", "s8a2q2", "s8a3q2", "s9eq1", "s1q1", "s1q2", "s9dq0", "s9dq1")

# Convert columns to factors for each data frame
rwa_exp_annual <- rwa_exp_annual %>% 
    mutate(across(any_of(factor_columns), as_factor))
rwa_exp_monthly <- rwa_exp_monthly %>% 
    mutate(across(any_of(factor_columns), as_factor))
rwa_exp_weekly <- rwa_exp_weekly %>% 
    mutate(across(any_of(factor_columns), as_factor))
rwa_exp_weekly2 <- rwa_exp_weekly2 %>% 
    mutate(across(any_of(factor_columns), as_factor))
rwa_exp_other <- rwa_exp_other %>% 
    mutate(across(any_of(factor_columns), as_factor))
rwa_individual <- rwa_individual %>% 
     mutate(across(any_of(factor_columns), as_factor))
rwa_inc <- rwa_inc %>% 
    mutate(across(any_of(factor_columns), as_factor))
    


```

## Descriptive statistics
```{r}
#Get household size for each household

hhsize <- rwa_individual %>% 
    group_by(hhid) %>% 
    summarise(hhsize = n())
```

```{r}

# Total population in Rwanda
rwa_individual %>% 
    summarise(sum(weight))
```
## Analysis
```{r}

# Select colums from individual survey file to merge with expenditure file

rwa_individual_age <- rwa_individual %>% 
  dplyr::select(hhid, weight, s1q1, s1q3y) %>%
  rename(pop_weight = weight, age = s1q3y) %>%
  mutate(age = as.numeric(age))

# Merge expenditure files and individual survey file with age information

rwa_merge <- rwa_exp_annual %>% 
  left_join(rwa_individual_age, by = c("hhid" = "hhid"))


```


### Total expenditure vs share spent on education for the entire population(14580 households)
```{r}
# Get the total expenditure of the population at different quantiles by region, province and district
# The local currency is then converted to 2017_ppp LCU and 2017_ppp LCU equivalent values in 2025
    # PPPs are both currency conversion factors and spatial price indexes. PPPs convert different currencies to a common currency and, in the process of conversion, equalize their purchasing power by controlling differences in price levels between countries.

tot_expenditure_region <- rwa_exp_annual %>% 
    group_by(region, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE)) %>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_expenditure_province <- rwa_exp_annual %>% 
    group_by(province, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_expenditure_district <- rwa_exp_annual %>% 
    group_by(district, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

```

```{r}
# Get the total expenditure of the population on education  at different quantiles by region, province and district
# The local currency is then converted to 2017_ppp LCU and 2017_ppp LCU equivalent values in 2025 
# Coicop code that start with 10 represents expenditure on education

tot_educ_expenditure_region <- rwa_exp_annual %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(region, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE)) %>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_educ_expenditure_province <- rwa_exp_annual %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(province, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_educ_expenditure_district <- rwa_exp_annual %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(district, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

```


```{r}
# What percentage of total expenditure is spent on education

# Merge the total expenditure and education expenditure data frames to enable us to get educ_share as a percentage of total education

rwa_educ_exp_share_region <- merge(tot_expenditure_region, tot_educ_expenditure_region, by = c("region", "quintile"))
rwa_educ_exp_share_province <- merge(tot_expenditure_province, tot_educ_expenditure_province, by = c("province", "quintile"))
rwa_educ_exp_share_district <- merge(tot_expenditure_district, tot_educ_expenditure_district, by = c("district", "quintile"))
```


```{r}
# Calculate education share as a total of total expenditure for the entire population by region, province and district

rwa_educ_exp_share_region <- rwa_educ_exp_share_region %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)

rwa_educ_exp_share_province <- rwa_educ_exp_share_province %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)

rwa_educ_exp_share_district <- rwa_educ_exp_share_district %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)

#Save the files as csv

rwa_educ_exp_share_district %>% 
    write_csv("rwa_educ_exp_district.csv")

rwa_educ_exp_share_region %>% 
    write_csv("rwa_educ_exp_region.csv")

rwa_educ_exp_share_province %>% 
    write_csv("rwa_educ_exp_province.csv")

```


```{r}
# On average what percent of total expenditure is spent on education by the entire population(0.16%)

tot_expenditure_pop <- rwa_exp_annual %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE)) %>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_expenditure_educ_pop <- rwa_exp_annual %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE)) %>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])




```


### Total expenditure vs share spent on education for households with children between 0-6(8137 households)

```{r}
# Get population that is less than 6 years old

rwa_early_childhood <- rwa_individual %>% 
  dplyr::select(hhid, weight, s1q1, s1q3y) %>%
  rename(pop_weight = weight, age = s1q3y) %>%
  mutate(age = as.numeric(age)) %>% 
  filter(age <= 6)

# What is the total population that is between 0-6(ecd)- 2,301,045 people

rwa_early_childhood %>% 
    summarise(pop_ecd = sum(pop_weight))
```


```{r}
# How many households have children between ages 0-6
#8137 households have children between 0-6

n_distinct(rwa_early_childhood$hhid)
unique(rwa_early_childhood$hhid)
```

```{r}
n_distinct(househld_ecd$hhid)
```


```{r}
# Get all households ids with children between ages 0 and 6

househld_ecd <- rwa_exp_annual %>% 
    filter(hhid %in% unique(rwa_early_childhood$hhid))
```

```{r}
# What is the population of all households with children between ages 0 and 6
# 7,871,518

rwa_pop_hh_ecd <- rwa_individual %>% 
  dplyr::select(hhid, weight, s1q1, s1q3y) %>%
  rename(pop_weight = weight, age = s1q3y) %>%
  mutate(age = as.numeric(age)) %>% 
  filter(hhid %in% unique(rwa_early_childhood$hhid))

rwa_pop_hh_ecd %>% 
    summarise(pop_hh_ecd = sum(pop_weight))
```



```{r}

# For all households with children between 0-6 get the total expenditure and the percentage spent on education

tot_expenditure_region_ecd <- househld_ecd %>% 
    group_by(region, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE)) %>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_expenditure_province_ecd <- househld_ecd %>% 
    group_by(province, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_expenditure_district_ecd <- househld_ecd %>% 
    group_by(district, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])
```



```{r}
# Get the total expenditure of households with children aged 0-6 on education  at different quantiles by region

# Coicop code that start with 10 represents expenditure on education

tot_educ_expenditure_region_ecd <- househld_ecd %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(region, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE)) %>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_educ_expenditure_province_ecd <- househld_ecd %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(province, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_educ_expenditure_district_ecd<- househld_ecd %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(district, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])
```


```{r}
# Merge different dataframes

rwa_educ_exp_share_region_ecd <- merge(tot_expenditure_region_ecd, tot_educ_expenditure_region_ecd, by = c("region", "quintile"))
rwa_educ_exp_share_province_ecd <- merge(tot_expenditure_province_ecd, tot_educ_expenditure_province_ecd, by = c("province", "quintile"))
rwa_educ_exp_share_district_ecd <- merge(tot_expenditure_district_ecd, tot_educ_expenditure_district_ecd, by = c("district", "quintile"))
```


```{r}
# Calculate education share as a total of total expenditure

rwa_educ_exp_share_region_ecd <- rwa_educ_exp_share_region_ecd %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)

rwa_educ_exp_share_province_ecd <- rwa_educ_exp_share_province_ecd %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)

rwa_educ_exp_share_district_ecd <- rwa_educ_exp_share_district_ecd %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)


#Save the files as csv

rwa_educ_exp_share_district_ecd %>% 
    write_csv("output/rwa_educ_exp_district_ecd.csv")

rwa_educ_exp_share_region_ecd %>% 
    write_csv("output/rwa_educ_exp_region_ecd.csv")

rwa_educ_exp_share_province_ecd %>% 
    write_csv("output/rwa_educ_exp_province_ecd.csv")
```



### Total expenditure vs share spent on education for households with children between 0-6 who have reported expenditure on education(29 households)

```{r}
# Get number of households that declare expenditure on education

househld_educ_exp <- rwa_exp_annual %>% 
    filter((str_starts(s8a1q1, "10")) & s8a1q2 == "Yes")

household_educ_exp_10 <- rwa_exp_annual %>% 
    filter(hhid %in% househld_educ_exp$hhid)
```


```{r}
##data check

n_distinct(househld_educ_exp$hhid)
n_distinct(household_educ_exp_10$hhid)
```


```{r}

# For all households with children between 0-6 who report spending on education, get the total expenditure 

tot_expenditure_region_hh10 <- household_educ_exp_10 %>% 
    group_by(region, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE)) %>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_expenditure_province_hh10 <- household_educ_exp_10 %>% 
    group_by(province, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_expenditure_district_hh10 <- household_educ_exp_10 %>% 
    group_by(district, quintile) %>% 
    summarise(exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(exp_ppp_per_international_dollar = exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

```


```{r}
# For all households with children between 0-6 who report spending on education, get  the amount spent on education

tot_educ_expenditure_region_hh10 <- household_educ_exp_10 %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(region, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE)) %>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_educ_expenditure_province_hh10 <- household_educ_exp_10 %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(province, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])

tot_educ_expenditure_district_hh10<- household_educ_exp_10 %>% 
    filter(str_starts(s8a1q1, "10")) %>% 
    group_by(district, quintile) %>% 
    summarise(educ_exp_2017 = sum(s8a1q3, na.rm = TRUE))%>% 
    mutate(educ_exp_ppp_per_international_dollar = educ_exp_2017/ ppp_2017$hhe_2017ppp_to_current_LCU[ppp_2017$ccode=="RWA" & ppp_2017$year==2025])
```


```{r}
# Merge the 2 dataframes so we can calculate %share on education

rwa_educ_exp_share_region_hh10 <- merge(tot_expenditure_region_hh10, tot_educ_expenditure_region_hh10, by = c("region", "quintile"))
rwa_educ_exp_share_province_hh10 <- merge(tot_expenditure_province_hh10, tot_educ_expenditure_province_hh10, by = c("province", "quintile"))
rwa_educ_exp_share_district_hh10 <- merge(tot_expenditure_district_hh10, tot_educ_expenditure_district_hh10, by = c("district", "quintile"))
```


```{r}
# Calculate education share as a total of total expenditure

rwa_educ_exp_share_region_hh10 <- rwa_educ_exp_share_region_hh10 %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)

rwa_educ_exp_share_province_hh10 <- rwa_educ_exp_share_province_hh10 %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)

rwa_educ_exp_share_district_hh10 <- rwa_educ_exp_share_district_hh10 %>% 
    mutate(educ_share = (educ_exp_ppp_per_international_dollar/exp_ppp_per_international_dollar)*100)


#Save the files as csv

rwa_educ_exp_share_district_hh10 %>% 
    write_csv("output/rwa_educ_exp_district_hh10.csv")

rwa_educ_exp_share_region_hh10 %>% 
    write_csv("output/rwa_educ_exp_region_hh10.csv")

rwa_educ_exp_share_province_hh10 %>% 
    write_csv("output/rwa_educ_exp_province_hh10.csv")
```

```{r}
rwa_educ_exp_share_province_hh10 %>% 
    summarise(xy = sum(educ_exp_2017))
```

```{r}

# Total population with education expenditure for households reporting a spending on education and with children between ages 0-6
pop_educ_exp <- rwa_individual %>% 
    dplyr::select(hhid, weight, s1q1, s1q3y) %>%
    rename(pop_weight = weight, age = s1q3y) %>%
    mutate(age = as.numeric(age)) %>%
    filter(hhid %in% (househld_educ_exp$hhid))
```


```{r}
# Total population with education expenditure for households reporting a spending on education and with children between ages 0-6

pop_educ_exp %>% 
    ungroup() %>% 
    summarise(pop = sum(pop_weight))

```
```{r}
# less than 6years

ecd <- pop_educ_exp %>% 
    filter(age<=6)


```

